Independent Component Analysis

### 2022-05-01 paper.tex
edited the algorithm section in paper

\section{Algorithms}
There are some detailed difference on how gradient algorithms are implemented, i.e. ordinary gradient descent or fixed point algorithm.
\subsection{objective function based on maximum likelihood  (Kannan)}

\subsection{FastICA}
\subsection{alternative objective function, skewness and kurtosis (Kannan)}
In this method, we maximize the non-Gaussianity given by
\begin{align}
J(Y) \propto \{E[G(Y)] - E[G(V)]\}^2,
\end{align}
where $Y$ is a random variable of zero mean and unit variance, which can be achieved by whitening the data, and $V$ is a Gaussian variable also of zero mean and unit variance. In function forms of $G_1$ and $G_2$ should not grow too fast to be robust. The following choice of $G$ is proposed by Hyvärinen and Oja.
\begin{align}
G_1(u) &= \frac{1}{a_1} \log (\cosh a_1 u). \\
G_1^{\prime}(u) &= \tanh(a_1 u).
\end{align}
We want to find the extrema of $E[G(Y)]$ to maximize $J(Y)$ under the constraints
\begin{align}
E[(w^T x)^2] = \| w\|^2 = 1.
\end{align}
From the Kuhn--Tucker conditions, the extremum condition is satisfied when
\begin{align}
E[x g(w^Tx)] - \beta w &= 0.
\end{align}
This equation can be solved by using Newton's method.
Let $F(w)$ denote the left-hand side of the equation, which we are trying to solve.
We update the value of $w$ in each iteration according to the following equation.
\begin{align}
w_{n+1} &= w_n - \frac{F(w_n)}{F^{\prime}(w_n)} \\
&= w_n - \frac{E[x g (w_n^T x)] - \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
&= \frac{w_n E[g^{\prime}(w_n^T x)] - \beta w_n - E[x g (w_n^T x)] + \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
&= \frac{w_n E[g^{\prime}(w_n^T x)] - E[x g (w_n^T x)]}{E[g^{\prime}(w_n^T x)] - \beta}.
\end{align}
\section{Example}
We implement FastICA according to the formalism described in the previous section.
The algorithm is applied to a following set of signal functions.
\begin{align}
s_1 &= \sin(2 t). \\
s_2 &= \mathrm{sign} [\sin(3 t)]. \\
s_3 &= \frac{1}{2\pi} (t~\mathrm{mod}~2\pi).
\end{align}
The mixing matrix is taken to be
\begin{align}
A &= \begin{bmatrix}
1 & 1 & 1 \\
0.5 & 2 & 1 \\
1.5 & 1 & 2
\end{bmatrix}.
\end{align}
The result from our implementation and FastICA are shown in Fig.~\ref{fig:sklearn}.
% \begin{figure}
% \includegraphics{scripts/sklearn.pdf}
% \caption{\label{fig:sklearn} FastICA results using our code and \lstinline{sklearn.decomposition.FastICA}}
% \end{figure}

\section{Applications}
3) example: sound single decomposition
4) EEG % (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
5) stock price

@@

\section{PCA vs. ICA}
% (https://www.cs.ubc.ca/~jnutini/documents/mlrg_pca.pdf)

>>>>

\section{Algorithms}
In this section, we will discuss some of the common algorithms based on the previous mathematical formalisms. We will categorize them into moment-based approaches and information-based approaches. In either case, the problem is equivalent to an optimization problem. There are some detailed differences on how this optimization problem is implemented, i.e. ordinary gradient descent, Newton's method or fixed point algorithm. We will simply specify what we implemented for various examples that we will discuss later. 
\subsection{Moment-based approach}
In this method, we maximize the non-Gaussianity given by
\begin{align}
J(Y) \propto \{E[G(Y)] - E[G(V)]\}^2,
\end{align}
where $Y$ is a random variable of zero mean and unit variance, which can be achieved by whitening the data, and $V$ is a Gaussian variable also of zero mean and unit variance. In function forms of $G_1$ and $G_2$ should not grow too fast to be robust. The following choice of $G$ is proposed by Hyvärinen and Oja.
\begin{align}
G_1(u) &= \frac{1}{a_1} \log (\cosh a_1 u). \\
G_1^{\prime}(u) &= \tanh(a_1 u).
\end{align}
We want to find the extrema of $E[G(Y)]$ to maximize $J(Y)$ under the constraints
\begin{align}
E[(w^T x)^2] = \| w\|^2 = 1.
\end{align}
From the Kuhn--Tucker conditions, the extremum condition is satisfied when
\begin{align}
E[x g(w^Tx)] - \beta w &= 0.
\end{align}
This equation can be solved by using Newton's method.
Let $F(w)$ denote the left-hand side of the equation, which we are trying to solve.
We update the value of $w$ in each iteration according to the following equation.
\begin{align}
w_{n+1} &= w_n - \frac{F(w_n)}{F^{\prime}(w_n)} \\
&= w_n - \frac{E[x g (w_n^T x)] - \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
&= \frac{w_n E[g^{\prime}(w_n^T x)] - \beta w_n - E[x g (w_n^T x)] + \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
&= \frac{w_n E[g^{\prime}(w_n^T x)] - E[x g (w_n^T x)]}{E[g^{\prime}(w_n^T x)] - \beta}.
\end{align}
\subsection{Information theoretic approach}

\section{Applications}
We implement FastICA according to the formalism described in the previous section.
The algorithm is applied to a following set of signal functions.
\begin{align}
s_1 &= \sin(2 t). \\
s_2 &= \mathrm{sign} [\sin(3 t)]. \\
s_3 &= \frac{1}{2\pi} (t~\mathrm{mod}~2\pi).
\end{align}
The mixing matrix is taken to be
\begin{align}
A &= \begin{bmatrix}
1 & 1 & 1 \\
0.5 & 2 & 1 \\
1.5 & 1 & 2
\end{bmatrix}.
\end{align}
The result from our implementation and FastICA are shown in Fig.~\ref{fig:sklearn}.
% \begin{figure}
% \includegraphics{scripts/sklearn.pdf}
% \caption{\label{fig:sklearn} FastICA results using our code and \lstinline{sklearn.decomposition.FastICA}}
% \end{figure}

%3) example: sound single decomposition
%4) EEG % (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
%5) stock price

@@
